---
title: "Technical Report"
author: "Evan Pugh"
date: "12/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(glmnet)
library(tidyverse)
library(broom)
library(glmnet)
library(ggplot2)
library(tree)
library(imager)
library(dplyr)
library(ISLR)
library(gbm)
OECD <- load.image("https://upload.wikimedia.org/wikipedia/commons/thumb/4/4c/OECD_member_states_map.svg/2560px-OECD_member_states_map.svg.png")

```

### Abstract

This project seeks to analyze what factors effect the death rate behind specific illnesses. The illnesses we have chosen to analyze are HIV-AIDs, Influenza, Tuberculosis, Neoplasms (tumors) and Pneumonia.

### Introduction

Our data comes from the OECD, or Organisation for Economic Co-operation and Development. The OECD is a group of nations who share data that they have gathered about themselves in order to seek solutions to shared problems, identify good policies and coordinate efforts between member countries.  Here is a map highlighting all 36 nations in the OECD:
```{r}
plot(OECD, axes = FALSE)

```
As you can probably tell from looking at the map, most of the countries on this map are countries that have a high income per capita and a high Human Development Index. In addition to these 36 countries, our data set also contains data on Brazil, Colombia, Costa Rica, Russia and South Africa. For this analysize we have choosen to measure data from 1990 to 2016


#### Exploratory Data Analysis

```{r}
death_rates <- read_csv("data/death_rates.csv")
employment <- read_csv("data/employment.csv")
GDPs <- read_csv("data/GDPs.csv")
insurance <- read_csv("data/insurance.csv")
populations <- read_csv("data/populations.csv")
PublicExp <- read_csv("data/PublicExp.csv")
GovtExp <- read_csv("data/GovtExp.csv")
employment = employment[employment$Measure=="Density per 1 000 population (head counts)",]
employment = cbind(employment[6],employment[8:9])

death_rates = cbind(death_rates[2],death_rates[6],death_rates[8:9])
GDPs = cbind(GDPs[6],GDPs[8:9])
populations = cbind(populations[6],populations[8:9])
insurance = cbind(insurance[6],insurance[8:9])
GovtExp = cbind(GovtExp[10],GovtExp[12],GovtExp[19])
PublicExp = cbind(PublicExp[10],PublicExp[12],PublicExp[19])

```

Let's start by plotting the number of deaths per 100,000 people for each of our five causes of death by year. Each of the 41 countries has been given a unique color and shape combination in order to indentify it and track it.

Units: 

All disease variables are reported as deaths per 100,000 people

Employment represents number of people whose job involves human health activities, residential care and social work activities per 1,000 people.

Population is the population of the country measured in thousands of people.

GDP is measured in US dollars per capita

Coverage represents the percentage of the population who are covered by public or private insurance.

GovtExp is the amount of money the government spends per person on healthcare costs with a constant PPP (also includes compulsory healthcare costs)

PublicExp is the amount of money individuals spend on healthcare per capita with a constant PPP.



```{r}
colors = rep(c("black","red","blue","green","violet","yellow","orange"),6)
shapes = rep(c(1,2,3,4,5,6),7)

death_rates=death_rates[death_rates$Year>1989,]
HIV = death_rates[death_rates$Variable == "HIV-AIDS",]
Flu = death_rates[death_rates$Variable == "Influenza",]
Tub = death_rates[death_rates$Variable == "Tuberculosis",]
Neo = death_rates[death_rates$Variable == "Neoplasms",]
Pne= death_rates[death_rates$Variable == "Pneumonia",]

ggplot(final, aes(x=Year, y=HIV,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from HIV-AIDs per 100,000 people")
  
ggplot(final, aes(x=Year, y=Influenza,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from Influenza per 100,000 people")

ggplot(final, aes(x=Year, y=Tuberculosis,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from Tuberculosis per 100,000 people")

ggplot(final, aes(x=Year, y=Neoplasms,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from Neoplasms per 100,000 people")

ggplot(final, aes(x=Year, y=Pneumonia,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from Pneumonia per 100,000 people")
```



```{r}
final = death_rates[,2:3]
final = unique(final)
final = left_join(final,HIV[,2:4], by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country","Year", "HIV-Aids")

final = left_join(final,Flu[,2:4], by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country","Year", "HIV-Aids","Influenza")

final = left_join(final,Tub[,2:4], by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country","Year", "HIV-Aids","Influenza","Tuberculosis")

final = left_join(final,Neo[,2:4], by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country","Year", "HIV-Aids","Influenza","Tuberculosis","Neoplasms")

final = left_join(final,Pne[,2:4], by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV-Aids", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia")

final = left_join(final,employment, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV-Aids", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment")

final = left_join(final,populations, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV-Aids", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment","Population")

final = left_join(final,GDPs, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment","Population", "GDP")

final = left_join(final,insurance, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment","Population", "GDP","Coverage")

final = left_join(final,GovtExp, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment","Population", "GDP","Coverage","GovtExp")

final = left_join(final,PublicExp, by = c("Country"="Country","Year"="Year"))
names(final) <- c("Country", "Year", "HIV", "Influenza", "Tuberculosis", "Neoplasms", "Pneumonia","Employment","Population", "GDP","Coverage","GovtExp","PublicExp")

final = unique(final)
```


#### Exercise 4

```{r}
mHIV = lm(final$HIV ~ final$Year + final$Employment + final$Coverage + final$PublicExp)
summary(mHIV)

mFlu = lm(final$Influenza ~ final$Year + final$Employment + final$Population + final$GDP)
summary(mFlu)

mTub = lm(final$Tuberculosis ~  final$Population + final$GDP + final$GovtExp+final$Coverage*final$Population+final$Coverage*final$Employment)
summary(mTub)

mPne = lm(final$Pneumonia ~ final$Year + final$Employment + final$Population + final$GDP + final$Coverage + final$GovtExp + final$PublicExp )
summary(mPne)

mNeo = lm(final$Neoplasms~ final$Year + final$Employment + final$Population + final$GDP + final$Coverage + final$GovtExp)
summary(mNeo)

```

So in summary:

Our model for HIV had an R-squared of 0.2264 and found that the Year, Employment, Coverage, and Public Expenditure were statistically significant variables.

Our model for Influenza had an R-squared of 0.2647 and found that the Year, Employment,Population and GDP were statistically significant variables (Public Expenditure just missed the cutoff).

Our model for Tuberculosis had an R-squared of 0.3871 and found that the Population, GovtExp, Coverage, Employment, the interaction term between population and coverage were statistically significant variables.

Our model for Pneumonia had an R-squared of 0.2719 and found that all of our variables were statistically significant variables.

Our model for Neoplasms had an R-squared of 0.3724 and found that the Year, Population, GDP Coverage, and Government Expenditure were statistically significant variables.

Now, undoubtally there are some variables effecting this data that are not in this data set. There are a tens of thousands of  variables that could effect these death rates, everything from the climate to the form of government. To try and account for many of these, let's add Country to the list of variables as a proxy for the variables that we have not accounted for but are mostly static for a given country.

```{r}

m2HIV = lm(final$HIV ~ final$Year + final$Population + final$GDP + final$Country)
summary(m2HIV)

m2Flu = lm(final$Influenza ~ final$Year + final$Population + final$GDP + final$GovtExp + final$PublicExp + final$Country)
summary(m2Flu)

m2Tub = lm(final$Tuberculosis ~ final$Year + final$Employment + final$Population + final$GDP + final$Coverage + final$GovtExp + final$PublicExp + final$Country)
summary(m2Tub)

m2Pne = lm(final$Pneumonia ~ final$Year + final$Employment + final$Population + final$GDP + final$Coverage + final$GovtExp + final$PublicExp + final$Country)
summary(m2Pne)

m2Neo = lm(final$Neoplasms~ final$Year + final$Employment + final$Population + final$GDP + final$Coverage + final$GovtExp + final$PublicExp + final$Country)
summary(m2Neo)
```
This model for HIV had an R-squared of 0.7999, an increase of 0.5456 from our previous model.

This model for Influenza had an R-squared of 0.4816, an increase of 0.1814 from our previous model.

This model for Tuberculosis had an R-squared of 0.8994, an increase of 0.5419 from our previous model.

This model for Pneumonia has an R-squared of 0.8126, an increase of 0.5407 from our previous model.

This model for Neoplasms had an R-squared of 0.9648, an increase of 0.5907 from our previous model.

So HIV, Tuberculosis, Pneumonia and Neoplasms are all heavily effected by variables that we do not have that are baked in to each of our countries, while Influenza is only mildly effected by these variables.


Finding optimal lambdas for glm
```{r}
set.seed(42)


testH = na.omit(final[,-4:-7])


x1 <- testH %>% select(Year, Employment, Population, GDP, Coverage, GovtExp, PublicExp) %>% data.matrix()
y1 <- testH$HIV
lambdas <- 10^seq(3, -10, by = -.1)
cv_fit <- cv.glmnet(x1, y1, alpha = .8, lambda = lambdas)

cv_fit$lambda.min
temp = final[,-3]
test2 = na.omit(temp[,-4:-6])


x2 <- test2 %>% select(Year, Employment, Population, GDP, Coverage, GovtExp, PublicExp) %>% data.matrix()
y2 <- test2$Influenza
lambdas <- 10^seq(1, -8, by = -.1)
cv_fit2 <- cv.glmnet(x2, y2, alpha = .8, lambda = lambdas)
cv_fit2$lambda.min
temp = final[,-3:-4]
test3 = na.omit(temp[,-4:-5])

x3 <- test3 %>% select(Year, Employment, Population, GDP, Coverage, GovtExp, PublicExp) %>% data.matrix()
y3 <- test3$Tuberculosis

cv_fit3 <- cv.glmnet(x3, y3, alpha = .8, lambda = lambdas)
cv_fit3$lambda.min

temp = final[,-3:-5]
test4 = na.omit(temp[,-4])

x4 <- test4 %>% select(Year, Employment, Population, GDP, Coverage, GovtExp, PublicExp) %>% data.matrix()
y4 <- test4$Neoplasms

cv_fit4 <- cv.glmnet(x4, y4, alpha = .8, lambda = lambdas)
cv_fit4$lambda.min

temp = final[,-3:-6]
test5 = na.omit(temp)

x5 <- test5 %>% select(Year, Employment, Population, GDP, Coverage, GovtExp, PublicExp) %>% data.matrix()
y5 <- test5$Pneumonia

cv_fit5 <- cv.glmnet(x5, y5, alpha = .8, lambda = lambdas)
cv_fit5$lambda.min

glmHIV <- glmnet(x1, y1, alpha = .8, lambda = 0.0158)

summary(glmHIV$beta)
coef(mHIV)
```

Messing around with training and test data, test data for lm is wonky
```{r}
set.seed(42)
train = sample(1:nrow(x1), nrow(x1)/2)


x1_train = x1[train,]
y1_train = y1[train]
x1_test = x1[-train,]
y1_test = y1[-train]



fit1 <- glmnet(x1_train, y1_train, alpha = .8, lambda = 0.0158)

preds <- predict(fit1,newx = x1_test)
actual <- y1_test
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq1 <- 1 - rss/tss
rsq1


x2_train = x2[train,]
y2_train = y2[train]
x2_test = x2[-train,]
y2_test = y2[-train]

fit2 <- glmnet(x2_train, y2_train, alpha = .8, lambda = 0.0158)

preds <- predict(fit2,newx = x2_test)
actual <- y2_test
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq2 <- 1 - rss/tss
rsq2



x3_train = x3[train,]
y3_train = y3[train]
x3_test = x3[-train,]
y3_test = y3[-train]

fit3 <- glmnet(x3_train, y3_train, alpha = .8, lambda = 0.0158)

preds <- predict(fit3,newx = x3_test)
actual <- y3_test
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq3 <- 1 - rss/tss
rsq3



x4_train = x4[train,]
y4_train = y4[train]
x4_test = x4[-train,]
y4_test = y4[-train]

fit4 <- glmnet(x4_train, y4_train, alpha = .8, lambda = 0.0158)

preds <- predict(fit4,newx = x4_test)
actual <- y4_test
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq <- 1 - rss/tss
rsq



x5_train = x5[train,]
y5_train = y5[train]
x5_test = x5[-train,]
y5_test = y5[-train]

fit5 <- glmnet(x5_train, y5_train, alpha = .8, lambda = 0.0158)

preds <- predict(fit5,newx = x5_test)
actual <- y5_test
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq5 <- 1 - rss/tss
rsq5


```



Messing around with PCA
```{r}

set.seed(42)

temp = na.omit(final)

dim(temp)
train = sample(1:nrow(temp), nrow(temp)/2)
temp.pca <- prcomp(na.omit(cbind(temp[,2],temp[8:13])))
summary(temp.pca)


temp$PC1 <- temp.pca$x[,1]
temp$PC2 <- temp.pca$x[,2]
temp$PC3 <- temp.pca$x[,3]
temp$PC4 <- temp.pca$x[,4]

predict
ggplot(temp, aes(x = PC1+PC2+PC3+PC4, y = HIV)) + geom_point()

```


Tree analysis
```{r}

set.seed(42)

temp = na.omit(final)
train = sample(1:nrow(temp), nrow(temp)/2)
temp.pca <- prcomp(na.omit(cbind(temp[,2],temp[8:13])))
tree.HIV = tree(HIV ~Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, temp[train,])

preds <- predict(tree.HIV, temp[-train,])
actual <- temp[-train,3]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq1 <- 1 - rss/tss
rsq1

tree.Flu = tree(Influenza ~Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, temp[train,])
preds <- predict(tree.Flu,temp[-train,])
actual <- temp[-train,4]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq2 <- 1 - rss/tss
rsq2

tree.Tub = tree(Tuberculosis ~Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, temp[train,])
preds <- predict(tree.Tub,temp[-train,])
actual <- temp[-train,5]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq3 <- 1 - rss/tss
rsq3

tree.Neo = tree(Neoplasms ~Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, temp[train,])
preds <- predict(tree.Neo,temp[-train,])
actual <- temp[-train,6]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq4 <- 1 - rss/tss
rsq4

tree.Pne = tree(Pneumonia ~Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, temp[train,])
preds <- predict(tree.Pne,temp[-train,])
actual <- temp[-train,7]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsq5 <- 1 - rss/tss
rsq5
  plot(tree.Neo)
text(tree.Neo)
plot(tree.Flu)
text(tree.Flu)
plot(tree.HIV)
text(tree.HIV)
plot(tree.Tub)
text(tree.Tub)
plot(tree.Pne)
text(tree.Pne)
```


Boosted Tree

```{r}

gbm1 <- gbm(HIV ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)


preds = predict(gbm1, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,3]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqb1 <- 1 - rss/tss
rsqb1


gbm2 <- gbm(Influenza ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, data = temp[train,], distribution="gaussian", n.trees=50, interaction.depth=2)


preds = predict(gbm2, temp[-train,], n.trees = 50, type = "response")
actual <- temp[-train,4]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqb2 <- 1 - rss/tss
rsqb2

gbm3 <- gbm(Tuberculosis ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)



preds = predict(gbm3, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,5]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqb3 <- 1 - rss/tss
rsqb3

gbm4 <- gbm(Neoplasms ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)



preds = predict(gbm4, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,6]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqb4 <- 1 - rss/tss
rsqb4

gbm5 <- gbm(Pneumonia ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbm5, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,7]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqb5 <- 1 - rss/tss
rsqb5

View(gbm5)
plot(gbm1)

```



Boosted Tree

```{r}
tempHIV = na.omit(HIV)
tempHIV = tempHIV
ggplot(tempHIV, aes(x=Year, y=gbm1,color=Country,shape=Country)) + geom_point()+
  scale_shape_manual(values=shapes[0:41])+
  scale_color_manual(values=colors[0:41]) + ylab("Deaths from Neoplasms per 100,000 people")

gbmf1 = 
```


Lagged variables

```{r}

HIVLag = lag(final$HIV)
FluLag = lag(final$Influenza)
TubLag = lag(final$Tuberculosis)
NeoLag = lag(final$Neoplasms)
PneLag = lag(final$Pneumonia)


final$HIVLag <- HIVLag
final$FluLag <- FluLag
final$TubLag <- TubLag
final$NeoLag <- NeoLag
final$PneLag <- PneLag

temp = na.omit(final)

temp = temp[temp$Year >1990,]
gbml1 <- gbm(HIV ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp + HIVLag, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbml1, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,3]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqbl1 <- 1 - rss/tss
rsqbl1
rsqb1


gbml2 <- gbm(Influenza ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp + FluLag, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbml2, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,4]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqbl2 <- 1 - rss/tss
rsqbl2
rsq2

gbml3 <- gbm(Tuberculosis ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp + TubLag, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbml3, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,5]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqbl3 <- 1 - rss/tss
rsqbl3
rsqb3

gbml4 <- gbm(Neoplasms ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp + TubLag, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbml4, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,6]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqbl4 <- 1 - rss/tss
rsqbl4
rsqb4

gbml5 <- gbm(Pneumonia ~ Year + Employment + Population + GDP + Coverage + GovtExp + PublicExp + PneLag, data = temp[train,], distribution="gaussian", n.trees=1000, interaction.depth=3)

preds = predict(gbml5, temp[-train,], n.trees = 1000, type = "response")
actual <- temp[-train,7]
rss <- sum((preds - actual) ^ 2)
tss <- sum((actual - mean(actual)) ^ 2)
rsqbl5 <- 1 - rss/tss
rsqbl5
rsqb5
```
Best Models
```{r}
mHIV = gbml1
mFlu = gbm2
mTB = gbml3
mNeo = gbml4
mPne = gbml5

rsqs <- matrix(c(0.203,0.252,0.262,0.325,0.276,0.346,0.450,0.127,0.673,0.509,0.537,0.438,0.729,0.847,0.737,0.906,0.405,0.924,0.871,0.896),ncol=4,byrow=FALSE)
colnames(rsqs) <- c("Elastic Net","Random Forest","Boosted Tree", "Boosted Tree w/ Lag")
rownames(rsqs) <- c("HIV-AIDs","Influenza", "Tuberculosis", "Neoplasms", "Pneumonia")
View(rsqs)
sd(na.omit(final$Influenza))
mean(na.omit(final$Influenza))

```

Overall we found that for every cause of death except for Influenza, the Elastic Net performed the worst, the Random Forest performed the third best, the Boosted Tree performed the second best and the Boosted Tree with a unit of lag performed the best. For Influenza, the Elastic Next still performed the worst, but the next worst model was the Boosted Tree with a unit of lag, followed by a regular Boosted Tree with a Random Forest being the best model at predicting Influenza. We believe this is because Influenza is significantly different in how it is spread from the other diseases on this list. Flu outbreaks only happen for about a six week period each year and isn't a significant threat outside of those six weeks. 


```{r}
Aus = temp[temp$Country=="Australia",]

prediction3 = predict(gbml2, Aus, n.trees = 1000, type = "response",)
prediction4 = predict(gbm2, Aus, n.trees = 50, type = "response")
prediction1 = predict(gbml1, Aus, n.trees = 1000, type = "response",)
prediction2 = predict(gbm1, Aus, n.trees = 1000, type = "response")

ggplot(Aus, aes(x = Year, y = Influenza)) +
     geom_point() +
  geom_line(aes(y = prediction4), color = "green", size = 1)+
  geom_line(aes(y = prediction3), color = "red", size = 1)

ggplot(Aus, aes(x = Year, y = HIV)) +
     geom_point() +
  geom_line(aes(y = prediction2), color = "green",size = 1) + geom_line(aes(y = prediction1), color = "red", size = 1)

predictionHIV = predict(gbml1, temp, n.trees = 1000, type = "response")
predictionFlu = predict(gbm2, temp, n.trees = 50, type = "response")
predictionTB = predict(gbml3, temp, n.trees = 1000, type = "response")
predictionNeo = predict(gbml4, temp, n.trees = 1000, type = "response")
predictionPne = predict(gbml5, temp, n.trees = 1000, type = "response")

ggplot(temp, aes(x = Year, y = HIV-predictionHIV)) + geom_point()

ggplot(temp, aes(x = Year, y = Influenza-predictionFlu)) + geom_point()

ggplot(temp, aes(x = Year, y = Tuberculosis-predictionTB)) + geom_point()

ggplot(temp, aes(x = Year, y = Neoplasms-predictionNeo)) + geom_point()

ggplot(temp, aes(x = Year, y = Pneumonia-predictionPne)) + geom_point()

```
